---
title: "Individual Assignment3 - Appendix"
output: word_document
---
# Data source:

<https://www.consumerfinance.gov/data-research/consumer-complaints/search/api/v1/?company=BANK%20OF%20AMERICA%2C%20NATIONAL%20ASSOCIATION&date_received_max=2021-03-15&date_received_min=2011-12-01&field=all&format=csv&lens=overview&no_aggs=true&product=Bank%20account%20or%20service&product=Checking%20or%20savings%20account&size=27104&trend_depth=5&trend_interval=month>


## Reference:

[1] https://www.statisticssolutions.com/assumptions-of-linear-regression

[2] "Discovering Statistics Using R" by Andy Field, Jeremy Miles, Zoe Field.


## Appendix -1
```{r echo=FALSE}
path<-'C:\\Users\\SALONI\\OneDrive - University of Waterloo\\MSCI 718\\i-3\\complaints-final.csv'
```

```{r echo=TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)
dataset <- read.csv(path)
```

## Appendix -2

```{r}
dataset<- dataset %>% filter(Company=="BANK OF AMERICA, NATIONAL ASSOCIATION")
dataset <- na.omit(dataset)

dataset<-dataset %>% separate(Date.received, c("Split_Month","Split_Day","Split_Year"),remove=FALSE)
dataset$Split_Year<- as.integer(dataset$Split_Year)
dataset$Split_Month <- as.integer(dataset$Split_Month)
dataset$Split_Year <- (dataset$Split_Year + 2000)

all_years<-c(2011,2012,2013,2014,2015,2016,2017,2018,2019,2020,2021)
all_months<-c(1,2,3,4,5,6,7,8,9,10,11,12)
all_quaters<-c(1,2,3,4)
```

## Appendix -3

```{r echo=FALSE}
#Month wise cases
Month_wise_cases <- NULL
time <- 0
for (year in all_years) {
  for (month in all_months) {
    complains <- count(dataset %>% filter(Split_Month==month & Split_Year==year))
    if(complains!=0){
      time <- time + 1
      Month_wise_cases <- rbind(Month_wise_cases, data.frame(year,month,time, complains))
    }
  }
}

Month_wise_cases$complains<- Month_wise_cases$n
Month_wise_cases$n <- NULL
Month_wise_cases
```


## Appendix -3

```{r echo=FALSE}
# yearwise cases
Year_wise_cases <- NULL

for (year in all_years) {
  
    complains <- count(dataset %>% filter( Split_Year==year))
    if(complains!=0){
         Year_wise_cases <- rbind(Year_wise_cases, data.frame(year, complains))
    }
  
}
Year_wise_cases
```

## Appendix -4
one important assumption of the linear regression is that the relationship between the outcome and predictor variables is linear and additive. Here quarter and no of complains are strongly & positively related. Correlation can be given as:
```{r echo=FALSE}
dataset<- dataset %>%  filter(Split_Year!=2012 | Split_Month !=3)
df <- NULL
no_of_quarter <- 0
for (year in all_years) {
  for (quarter in all_quaters) {
    if(quarter==1){
       m1=1;m2=2;m3=3;
    }
    if(quarter==2){
       m1=4;m2=5;m3=6;
    }
    if(quarter==3){
       m1=7;m2=8;m3=9;
    }
    if(quarter==4){
       m1=10;m2=11;m3=12;
    }
    complains <- count(dataset %>% filter((Split_Month==m1| Split_Month==m2|Split_Month==m3)& Split_Year==year))
    if(complains!=0){
      no_of_quarter <- no_of_quarter + 1
      df <- rbind(df, data.frame(year,quarter,no_of_quarter, complains))
    }
  }
}
df$complains<- df$n
df$n <- NULL
#df
```

```{r}
cor( df$no_of_quarter,df$complains)*100

```

## Appendix -5

```{r echo=FALSE}
regressor = lm(formula = complains ~ no_of_quarter,   data = df)
summary(regressor)
```

## Appendix -6

```{r echo=FALSE}
confint(regressor)
```

## Appendix -7

```{r}
sigma(regressor)*100/mean(df$complains)
```



## Appendix -8

```{r echo=FALSE}
scatter.smooth(x=Month_wise_cases$time, y=Month_wise_cases$complains, main="complains ~ time")  # scatterplot
```

## Appendix -9

```{r echo=FALSE}
plot(regressor)
```

## Appendix -10

```{r echo=FALSE}
# assumption of homogenity of variance
std.residuals <- (regressor$residuals-mean(regressor$residuals))/sd(regressor$residuals)
std.fitted <- (regressor$fitted.values-mean(regressor$fitted.values))/sd(regressor$fitted.values)
plot(std.residuals,std.fitted)

```



## Appendix -11

```{r echo=FALSE}

regressor$residuals<-resid(regressor)
regressor$standardized.residuals<- rstandard(regressor)
regressor$studentized.residuals<-rstudent(regressor)
regressor$cooks.distance<-cooks.distance(regressor)
regressor$dfbeta<-dfbeta(regressor)
regressor$dffit<-dffits(regressor)
regressor$leverage<-hatvalues(regressor)
regressor$covariance.ratios<-covratio(regressor)

regressor$large.residual <- regressor$standardized.residuals > 2 | regressor$standardized.residuals < -2


sum(regressor$large.residual) 

#regressor[regressor$large.residual,c("sales", "airplay", "attract", "adverts","standardized.residuals")]
#regressor[regressor$large.residual , c("cooks.distance", "leverage", "covariance.ratios")]

regressor$fitted <- regressor$fitted.values

```

## Appendix -12

Assumption of normality
```{r echo=FALSE}


histogram<-ggplot(regressor, aes(regressor$studentized.residuals)) + 
  theme(legend.position ="none") + 
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  labs(x = "Studentized Residual", y = "Density")

```

```{r echo=FALSE}
histogram +  stat_function(fun = dnorm, args = list(mean = mean(regressor$studentized.residuals, na.rm = TRUE), 
                                         sd = sd(regressor$studentized.residuals, na.rm = TRUE)),
  colour= "red", size = 1)

```
## Appendix -13

```{r echo=FALSE}
qqplot.resid <- qplot(sample = regressor$studentized.residuals, stat="qq") + 
  labs(x ="Theoretical Values", y = "Observed Values") 
qqplot.resid

```
## Appendix -14

```{r echo=FALSE}
scatter <- ggplot(regressor, aes(regressor$fitted, regressor$studentized.residuals))
scatter + geom_point() + 
  geom_smooth(method = "lm", colour = "Blue")+ 
labs(x = "Fitted Values", y = "Studentized Residual")

```



## Appendix -15

```{r echo=FALSE}
init<-max(df$no_of_quarter)
newValues=tibble(no_of_quarter=c(init+1,init+2,init+3,init+4,init+5,init+6,init+7))
future_prediction<- predict.lm(regressor, newValues)
future_prediction
```

## Appendix -16

```{r echo=FALSE}
val_2022 <- sum(future_prediction[4:7])
val_2022
```


## Appendix -17

```{r echo=FALSE}
library(car)

durbinWatsonTest(regressor)
```

## Appendix -18

```{r echo=FALSE}
df<-df

fit <- lm(formula = complains ~ no_of_quarter,df)  
#install.packages("boot")
library(boot)
bs=function(formula,data,indices){
  d=data[indices,]
  fit=lm(formula,data=d)
  (coef(fit))
}  
results=boot(data=df,statistic=bs, R=2000,formula= complains ~ no_of_quarter)  
boot.ci(results,type="bca",index=1)  
boot.ci(results,type="bca",index=2)


```
