---
title: "Individual Assignment4- Appendix"
output: word_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr);library(tidyr);library(ggplot2);
library(boot); library(car); library(QuantPsyc);
library(ggfortify);
dataset <- read.csv('melb_data.csv')

```

## Appendix - 1
For latitude between -37.6 and -38.0 there are many houses,which would mean that it’s an ideal residential location.

```{r echo=FALSE}
scatter1 <- ggplot(dataset, aes(Lattitude,Propertycount))
scatter1 + geom_jitter()

```


But when we talk about longitude we can see that concentration 
is high between 144.8 to 145.1 (This is close to the location of CBD)
```{r echo=FALSE}

scatter2 <- ggplot(dataset, aes(Longtitude,Propertycount ))
scatter2 + geom_jitter()
```


Given number of rooms between 3-5, there seems to be houses in all 
price ranges. As the number of rooms increases we can find higher 
price range houses. For 1-2 room houses there are lower price range houses

```{r echo=FALSE}
scatter3 <- ggplot(dataset, aes(Rooms,Price ))
scatter3 + geom_jitter()
```


Closer to CBD, there are a number of houses with varying price range 
Farther away, the number reduces drastically and the price too

```{r echo=FALSE}
scatter4 <- ggplot(dataset, aes(Distance,Price ))
scatter4 + geom_point()
```


As the number of bathrooms becomes more than 4, the number of houses 
decreases drastically. 


```{r echo=FALSE}
scatter5 <- ggplot(dataset, aes(Bathroom,Propertycount ))
scatter5 + geom_jitter()
```



Initially, the price range increases as the number of bathrooms increases.

```{r echo=FALSE}
scatter6 <- ggplot(dataset, aes(Bathroom,Price ))
scatter6 + geom_jitter()

```

People with 2 or 4 cars have been seen to be buying the costliest of houses 
As people have more number of houses the price decreases

```{r echo=FALSE}
scatter7 <- ggplot(dataset, aes(Car,Price ))
scatter7 + geom_jitter()
```


```{r echo=FALSE}
scatter7 <- ggplot(dataset, aes(Car,Propertycount ))
scatter7 + geom_jitter()
```

Landsize doesn't distinctly affect the price of a house

```{r echo=FALSE}
scatter8 <- ggplot(dataset, aes(Landsize,Price ))
scatter8 + geom_jitter()
```


As building area increases we can see a slight shift of the range of the houses available

```{r echo=FALSE}
scatter9 <- ggplot(dataset, aes(BuildingArea,Propertycount ))
scatter9 + geom_jitter()
```




## Appendix - 2

```{r echo=FALSE}

col <- ggplot(dataset, aes(Propertycount ,Type))
col + geom_col() 


```
## Appendix - 3

```{r echo=FALSE}
col <- ggplot(dataset, aes(Price ,Type))
col + geom_col() 
```

## Appendix - 4

```{r echo=FALSE}
dataset<-na.omit(dataset)

dataset$Type[dataset$Type=='h']<-'House'
dataset$Type[dataset$Type=='t']<-'NonHouse'
dataset$Type[dataset$Type=='u']<-'NonHouse'



dataset <- dataset %>%  filter (Suburb=='Reservoir'|Suburb=='Richmond'| Suburb=='Preston'|Suburb=='Brunswick'|Suburb=='Bentleigh East')
```

## Appendix -4b

```{r echo=FALSE}

dataset_new<- dataset %>% select_if(is.numeric)

library(corrplot)
res <- cor(dataset_new)
#round(res, 2)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

```

## Appendix -5

```{r echo=FALSE}
model.1<-lm(formula = Price ~ BuildingArea+Postcode+Rooms+Bathroom+Distance,   data = dataset,na.action=na.exclude)
summary(model.1)
```



```{r echo=FALSE}
autoplot(model.1, which = 1:6, label.size = 1)
```



```{r echo=FALSE}
model.2<-lm(formula = Price ~ BuildingArea+Postcode+Rooms+Bathroom+Distance+Type,data = dataset,na.action=na.exclude)
summary(model.2)
```



```{r}

autoplot(model.2, which = 1:6, label.size = 1,  colour = 'Type')

```

```{r echo=FALSE}

dataset$Price<- log(dataset$Price)

dataset$Postcode<- log(dataset$Postcode)
dataset$Rooms <- log(dataset$Rooms)
dataset$Bathroom <- log(dataset$Bathroom)
dataset$Distance <- log(dataset$Distance+1)
dataset$BuildingArea <- log(dataset$BuildingArea)
model.1<-lm(formula = Price ~ BuildingArea+Postcode+Rooms+Bathroom+Distance,   data = dataset,na.action=na.exclude)
model.2<-lm(formula = Price ~ BuildingArea+Postcode+Rooms+Bathroom+Distance+Type,   data = dataset,na.action=na.exclude)
```



## Appendix - 6
```{r}
summary(model.1)
```


```{r}
summary(model.2)
```

```{r}
lm.beta(model.1)
```


```{r}
lm.beta(model.2)
```


## Appendix -7

```{r}
dataset$residuals<-resid(model.1)  #page 323
dataset$standardized.residuals<- rstandard(model.1) #page 324
dataset$studentized.residuals<-rstudent(model.1)
dataset$cooks.distance<-cooks.distance(model.1)
dataset$dfbeta<-dfbeta(model.1)
dataset$dffit<-dffits(model.1)
dataset$leverage<-hatvalues(model.1)
dataset$covariance.ratios<-covratio(model.1)


e <- ggplot(dataset, aes(y=standardized.residuals, x=leverage))
e+geom_point()

```

```{r}
summary(dataset$cooks)
```

```{r}
dataset$residuals<-resid(model.2)  #page 323
dataset$standardized.residuals<- rstandard(model.2) #page 324
dataset$studentized.residuals<-rstudent(model.2)
dataset$cooks.distance<-cooks.distance(model.2)
dataset$dfbeta<-dfbeta(model.2)
dataset$dffit<-dffits(model.2)
dataset$leverage<-hatvalues(model.2)
dataset$covariance.ratios<-covratio(model.2)


e <- ggplot(dataset, aes(y=standardized.residuals, x=leverage))
e+geom_point()

```
```{r}
summary(dataset$cooks)
```

## Appendix-8

```{r}
vif(model.1)
1/vif(model.1)
mean(vif(model.1))
```

```{r}
vif(model.2)
1/vif(model.2)
mean(vif(model.2))
```

## Appendix -9 

```{r}
confint(model.2)
```

```{r}
exp(confint(model.2))
```

## Appendix -10

```{r}
anova(model.1,model.2)
```



## Appendix -11

### Price prediction

```{r echo=FALSE}
#builting area 120 to 1041
#Postcode  3165    3056  3072
#Rooms 3 2
#Bathroom 2
#Distance 12
#BuildingArea+Postcode+Rooms+Bathroom+Distance+Type
newdata = data.frame(BuildingArea=100,Postcode=3056,Rooms=3,Bathroom=2,Distance=12)
log_newdata<-log(newdata)
log_price=predict(model.1, log_newdata) 
log_price
```
```{r}
exp(log_price)
```
```{r}
log_newdata
```

```{r}
log_newdata$Type='House'
```

```{r }

log_price=predict(model.2, log_newdata) 
exp(log_price)
```

```{r}
log_newdata$Type='NonHouse'
```

```{r}

log_price=predict(model.2, log_newdata) 
exp(log_price)
```
